---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


**Warning: tidyABS is in early development. Please use for experimentation only.**

## Overview

The tidyABS package converts ABS excel tables to tidy data frames. It uses rules-of-thumb to determine the structure of excel tables, however it sometimes requires pointers from the user. 


_Note: tidyABS works with .xlsx files only._

## Installation

The tidyABS package is not available on CRAN. It can be installed from github with the following script:
 
```{r eval = FALSE}
# install.packages("devtools")
devtools::install_github("ianmoran11/tidyABS")
```

## Usage

```{r include = FALSE}
options(tibble.print_min = 3, tibble.print_max = 3)
library(magick)
```

```{r message=FALSE, warning=FALSE}
library(tidyABS)
library(tidyverse)
```

Below is a short demonstration that tidies a table from the Australian Industry publication (catalogue 8155.0). 
```{r message=FALSE, warning=FALSE}

tidy_aus_industry_df <-
  tidyABS_example("australian-industry.xlsx") %>%
  process_sheet(sheets = "Table_1") %>%
  assemble_table_components()

tidy_aus_industry_df %>% str()
```

### Examples

The tidyABS package contains several example files. Use the helper `tidyABS_example()` function with no arguments to list these files:

```{r message=FALSE, warning=FALSE}
tidyABS_example()
```

#### Example 1: Australian Industry

```{r echo=FALSE, message=FALSE, warning=FALSE}
magick::image_read("australian-industry_table-1.png")
```

Above is the  first sheet of an excel workbook from the Australian Industry publication. We can retrieve the path of this file  using the `tidyABS_example` function:

```{r message=FALSE, warning=FALSE}
ai_path <- tidyABS_example("australian-industry.xlsx")
```

To process the sheet above, we pass the workbook file path to the `process_ABS_sheet` function and identify the sheet we'd like to tidy. 

```{r message=FALSE, warning=FALSE}
ai_processed <- process_sheet(path = ai_path, sheets = "Table_1")

ai_processed %>% str(1)
```

This produces a list of three data frames. They store the location and format information of row names (`row_groups`), column names (`col_groups`) and table data (`tabledata`).  

We can inspect the row and column names in  `ai_processed` using the `inspect_table_components` function.

```{r message=FALSE, warning=FALSE}
inspect_table_components(ai_processed)
```

We can also visualize how these groups are spatially layed out in the spreadsheet and how tidyABS will relate them to table values with the function `plot_table_components`. Row names directly to the left of their data points should be labelled "W", and column names directly above should be labelled "N". (See the [unpivotr](https://github.com/nacnudus/unpivotr) package for more information.)

```{r fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
plot_table_components(ai_processed) +
  ylim(-30, 0)
```

Finally, we can assemble the components into a tidy data frame using `assemble_table_components`.

```{r message=FALSE, warning=FALSE}
assemble_table_components(ai_processed) %>%
  glimpse()
```

### Example 2: Environmental-Economic Accounts

```{r echo=FALSE, message=FALSE, warning=FALSE}
magick::image_read("environmental-economic-accounts_table-6-01.png")
```

Here's an example that requires some manual work, the Environmental-Economic Accounts. Let's retrieve the path of our example workbook and proces `Table 6.1`:

```{r message=FALSE, warning=FALSE}
eea_path <- tidyABS_example("environmental-economic-accounts.xlsx")

eea_processed <- process_sheet(path = eea_path, sheets = "Table 6.1")
```

On visual inspection, we can see `row_group_01` has been given a "W" orientation to the data, not "WNW". 

```{r fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
plot_table_components(eea_processed)
```

Luckily, we can fix this with the `change_direction` function. 

```{r message=FALSE, warning=FALSE}
eea_processed <-
  eea_processed %>%
  change_direction("row_group_01", "WNW")
```

Plotting the table confirms the direction has been corrected. 

```{r fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
plot_table_components(eea_processed)
```

Finally, we can assemble the components into a tidy dataframe using `assemble_table_components`.

```{r message=FALSE, warning=FALSE}
assemble_table_components(eea_processed) %>%
  glimpse()
```

### Example 3: Consumer Price Index (time series)

```{r echo=FALSE, message=FALSE, warning=FALSE}
magick::image_read("consumer-price-index_data01.png")
```

Time series data require the user to manually identify the inner table cells. This is because some of the column names are numberic --- for example, collection month. 

I recommend using the [`readABS`](https://github.com/MattCowgill/readabs) package for this. It was created for importing ABS time series data and does not require manual identifcation of table cells. 

That said, here's how you would process this table with `tidyABS`.

```{r message=FALSE, warning=FALSE}
cpi_path <- tidyABS_example("consumer-price-index.xlsx")
```

We need to identify the inner table cells using the `manual_value_references` argument.This argument takes a vector of addresses, identifying the inner corners of the table. 

```{r message=FALSE, warning=FALSE}
cpi_processed <-
  process_sheet(
    path = cpi_path, sheets = "Data1",
    manual_value_references = "B11:AB292"
  )
```

Here is the resulting data frame. 

```{r message=FALSE, warning=FALSE}
assemble_table_components(cpi_processed) %>%
  glimpse()
```
