---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r options, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(tidyABS)
```

## Tidying excel sheets from The Survey of Earned Doctorates

Here are a couple of examples using tidyABS on a non-ABS excel table. 

#### Table 12. Doctorate recipients, by major field of study: Selected years, 1987–2017

##### Minimal code 
Here's a mininal example.
```{r message=FALSE, warning=FALSE}
tidyABS_example("PhD_major-field.xlsx") %>%
  process_ABS_sheet(path = ., sheets = 1, manual_value_references = c("B6", "B50", "O50", "O6")) %>%
  change_direction("row_group_01", "WNW") %>%
  change_direction("row_group_02", "WNW") %>%
  assemble_table_components() %>%
  glimpse()
```

##### Step-through

First, read in the sheet, specifying table corners.  

```{r message=FALSE, warning=FALSE}
phd_field_df_components <-
  tidyABS_example("PhD_major-field.xlsx") %>%
  process_ABS_sheet(path = ., sheets = 1, manual_value_references = c("B6", "B50", "O50", "O6"))
```

Check the orietation of cells and make corrections. 

```{r fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
phd_field_df_components %>% plot_table_components()

phd_field_df_components <-
  phd_field_df_components %>%
  change_direction("row_group_01", "WNW") %>%
  change_direction("row_group_02", "WNW")
```

Check the orietation of cells and make corrections. 
```{r message=FALSE, warning=FALSE}
phd_field_df <-
  phd_field_df_components %>%
  assemble_table_components()

phd_field_df %>% glimpse()
```

#### Table 22. Doctorate recipients, by subfield of study, citizenship status, ethnicity, and race: 2017

##### Minimal code 
Here's a mininal example.
```{r message=FALSE, warning=FALSE}
tidyABS_example("PhD_ subfield-citizenship-status-ethnicity-race.xlsx") %>%
  process_ABS_sheet(path = ., sheets = 1, manual_value_references = c("B7", "L7", "B277", "L277")) %>%
  change_direction("row_group_01", "WNW") %>%
  change_direction("row_group_02", "WNW") %>%
  change_direction("row_group_03", "WNW") %>%
  assemble_table_components() %>%
  glimpse()
```

##### Step-through

First, read in the sheet, specifying table corners.  

```{r message=FALSE, warning=FALSE}
phd_background_components <-
  tidyABS_example("PhD_ subfield-citizenship-status-ethnicity-race.xlsx") %>%
  process_ABS_sheet(path = ., sheets = 1, manual_value_references = c("B7", "L7", "B277", "L277"))
```

Check the orietation of cells and make corrections. 

```{r fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
phd_background_components %>% plot_table_components() + ylim(-30, 0)

phd_background_components <-
  phd_background_components %>%
  change_direction("row_group_01", "WNW") %>%
  change_direction("row_group_02", "WNW") %>%
  change_direction("row_group_03", "WNW")
```

Check the orietation of cells and make corrections. 
```{r message=FALSE, warning=FALSE}
phd_background_df <-
  phd_background_components %>%
  assemble_table_components()
```
