    
---
title: "Tidying excel sheets from The Survey of Earned Doctorates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extending ggplot2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 7, fig.align = "center")
library(ggplot2)
```


```{r options, message=FALSE, warning=FALSE, include=FALSE}
library(magick)
library(tidyverse)
library(tidyABS)
```

## Tidying excel sheets from The Survey of Earned Doctorates

Here are a couple of examples using tidyABS on a non-ABS excel table. 

#### Table 12. Doctorate recipients, by major field of study: Selected years, 1987–2017

##### Minimal code 
Here's a mininal example.
```{r message=FALSE, warning=FALSE}
tidyABS_example("PhD_major-field.xlsx") %>%
  process_sheet(path = ., sheets = 1, manual_value_references = "B6:O50") %>%
  change_direction("row_group_01", "WNW") %>%
  change_direction("row_group_02", "WNW") %>%
  assemble_table_components() %>%
  str()
```

##### Step-through

First, read in the sheet, specifying table corners.  

```{r message=FALSE, warning=FALSE}
phd_field_df_components <-
  tidyABS_example("PhD_major-field.xlsx") %>%
  process_sheet(path = ., sheets = 1, manual_value_references = "B6:O50")
```

Check the orietation of cells and make corrections. 

```{r fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
phd_field_df_components %>% plot_table_components()

phd_field_df_components <-
  phd_field_df_components %>%
  change_direction("row_group_01", "WNW") %>%
  change_direction("row_group_02", "WNW")
```

Check the structure of the table implied by `phd_field_df_components` against the original excel table. If tidyABS has not identified row and column names correctly, the output below will not line up with the original excel table.   

```{r fig.height=6, message=FALSE, warning=FALSE}
phd_field_df_components %>% reconstruct_table() 
```

Using `View()` is useful here:

```{r echo=FALSE, message=FALSE, warning=FALSE}
magick::image_read("phd_field_df_components_view.png")
```



Assemble.
```{r message=FALSE, warning=FALSE}
phd_field_df <-
  phd_field_df_components %>%
  assemble_table_components()

phd_field_df %>% glimpse()
```

#### Table 22. Doctorate recipients, by subfield of study, citizenship status, ethnicity, and race: 2017

##### Minimal code 
Here's a mininal example.
```{r message=FALSE, warning=FALSE}
tidyABS_example("PhD_ subfield-citizenship-status-ethnicity-race.xlsx") %>%
  process_sheet(path = ., sheets = 1, manual_value_references = "B7:L277") %>%
  change_direction("row_group_01", "WNW") %>%
  change_direction("row_group_02", "WNW") %>%
  change_direction("row_group_03", "WNW") %>%
  change_direction("row_group_04", "WNW") %>%
  assemble_table_components() %>%
  glimpse()
```

##### Step-through

First, read in the sheet, specifying table corners.  

```{r message=FALSE, warning=FALSE}
phd_background_components <-
  tidyABS_example("PhD_ subfield-citizenship-status-ethnicity-race.xlsx") %>%
  process_sheet(path = ., sheets = 1, manual_value_references = "B7:L277")

```

Check the orietation of cells and make corrections. 

```{r fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
phd_background_components %>% plot_table_components() + ylim(-30, 0)

phd_background_components <-
  phd_background_components %>%
  change_direction("row_group_01", "WNW") %>%
  change_direction("row_group_02", "WNW") %>%
  change_direction("row_group_03", "WNW") %>% 
  change_direction("row_group_04", "WNW")

```

Check the structure of the table implied by `phd_field_df_components` against the original excel table. If tidyABS has not identified row and column names correctly, the output below will not line up with the original excel table.   

```{r fig.height=6, message=FALSE, warning=FALSE}
phd_background_components %>% reconstruct_table()  
```

Using `View()` is useful here:

```{r echo=FALSE, message=FALSE, warning=FALSE}
magick::image_read("phd_background_components_view.png")
```


Assemble.
```{r message=FALSE, warning=FALSE}
phd_background_df <-
  phd_background_components %>%
  assemble_table_components()


phd_background_df %>% glimpse()
  
```
